---
title: "viz_and_eda_1"
author: "Lorraine Kwok"
date: "2019-09-25"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
# will only be using one function in ggridges
```

## create the weather data 

```{r load_data, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

## create a ggplot

```{r}
## ggplot(dataset of interest, define what's on the x and y axes) + type of plot
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

alternate way of making this plot

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

saving initial plots

```{r}
scatterplot =
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

scatterplot
```

adding color

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .4)
## color is map feature we're interested in and name is the variable of interest
## alpha blending adds transparency to plot, 0 = transparent, 1 is not at all 
```

why do 'aes' positions matter?

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .4) +
  geom_smooth(se = FALSE)
## geom_smooth is often used for exploratory analysis
```

vs

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .4) +
  geom_smooth(se = FALSE)

## resulting plot has individual points and curves color-coded by the name 
```

## create facets
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .4) +
  geom_smooth(se = FALSE) +
  facet_grid(~name)
## facet_grid(~name) tells R you want name in the columns
```

## create plot to analyze seasonal trend
```{r}
weather_df %>%
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(size = 2, se = FALSE) +
  facet_grid(~name)
## size = prcp -> the size of the datapoints are determined by the precipitation data
```

## what happens when you remove geom_point from code

```{r}
weather_df %>%
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth(size = 2, se = FALSE)
## se = FALSE removes the standard errors from the plot
```

## 2d density plot
```{r}
## install.packages("hexbin")
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex() 
```

## More kinds of plots 

```{r univariate_plot}
weather_df %>%
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() +
  facet_grid(~name)
## gives information on univariate distributio of tmax
## color = color of bar's outline
## fill = color of bar's fill
## position = "dodge" "unstacks" the bars
```

density plot - conveys similar information as histograms, but the plots are a bit more smoothed out

```{r}
weather_df %>%
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = .3) 
```

boxplot

```{r}
weather_df %>%
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
## each location get its own boxplot 
```

violin plot 

```{r}
weather_df %>%
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()
```

ridge plots
We're looking at the density of tmax for Waterhole, Waikiki and Central Park.

```{r}
weather_df %>%
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()
```

## saving a plot

```{r}
ggp_ridge_temp =
  weather_df %>%
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()

ggsave("ggplot_temp_ridge.pdf", ggp_ridge_temp)
## ggsave(file path name to be saved into, the name of the plot you want to save)
```


## embedding plots into R markdown 

```{r, fig.width = 6}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .4) +
  geom_smooth(se = FALSE)
```

```{r, fig.width = 12}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .4) +
  geom_smooth(se = FALSE)
```
