---
title: "Visualization and EDA Part 2"
author: "Lorraine Kwok"
data: "10/1/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
```

## create the weather data 

```{r load_data, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

## making new plots

start with an old plot

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)
```

add labels: 

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NDAA via rnoaa pacakge"
  )
```

x axis tick marks etc

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NDAA via rnoaa pacakge"
  ) +
  scale_x_continuous(
    breaks = c(-15, -5, 20),
    labels = c("-15C", "-5C", "20C")
  ) +
  scale_y_continuous(
    trans = "sqrt"
  )
# can also use scale_y_sqrt instead of scale_y_continuous(trans = "sqrt")
```

## setting colors 

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NDAA via rnoaa pacakge"
  ) + 
  scale_color_hue(
    name = "Weather Station",
    h = c(50, 250) 
  )

# name is a function to change the name of the legend
```

## viridis pacakage includes color palettes that are not built into base R 

```{r}
ggp_base = weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NDAA via rnoaa pacakge"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

ggp_base
```

## themes
are global things that tell you what the plot is 
often has to do with organizational things
overarching themes go first and any tweaks you want to make go after

change legend position 
```{r}
# legend is at the bottom 
ggp_base + 
  theme(legend.position = "bottom") 
```

can change background theme
```{r}
# black and white theme
ggp_base + 
  theme_bw() +
  theme(legend.position = "bottom") 

# minimum theme
ggp_base +
  theme_minimal() +
  theme(legend.position = "bottom") 
```

