---
title: "Visualization and EDA Part 2"
author: "Lorraine Kwok"
data: "10/1/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
library(patchwork)
```

## create the weather data 

```{r load_data, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

## making new plots

start with an old plot

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)
```

add labels: 

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NDAA via rnoaa pacakge"
  )
```

x axis tick marks etc

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NDAA via rnoaa pacakge"
  ) +
  scale_x_continuous(
    breaks = c(-15, -5, 20),
    labels = c("-15C", "-5C", "20C")
  ) +
  scale_y_continuous(
    trans = "sqrt"
  )
# can also use scale_y_sqrt instead of scale_y_continuous(trans = "sqrt")
```

## setting colors 

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NDAA via rnoaa pacakge"
  ) + 
  scale_color_hue(
    name = "Weather Station",
    h = c(50, 250) 
  )

# name is a function to change the name of the legend
```

## viridis pacakage includes color palettes that are not built into base R 

```{r}
ggp_base = weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NDAA via rnoaa pacakge"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

ggp_base
```

## themes
are global things that tell you what the plot is 
often has to do with organizational things
overarching themes go first and any tweaks you want to make go after

change legend position 
```{r}
# legend is at the bottom 
ggp_base + 
  theme(legend.position = "bottom") 
```

can change background theme
```{r}
# black and white theme
ggp_base + 
  theme_bw() +
  theme(legend.position = "bottom") 

# minimum theme
ggp_base +
  theme_minimal() +
  theme(legend.position = "bottom") 
```

answering question about where the warning messages come from 
```{r}
weather_df %>%
  filter(is.na(tmin))
# ggplot is removing those 15 data points with missing x-values because it can't plot those data points 
```

can set theme at the very beginning to ensure that the plots always look the same
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))

## creating plots with more than one dataset 

```{r}
central_park =
  weather_df %>%
  filter(name == "CentralPark_NY")

waikiki =
  weather_df %>%
  filter(name == "Waikiki_HA")

# create scatterplot with waikiki dataset and added a line using central park dataset

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park)
```

(brief aside about colors)

```{r}
# can specify color in geom_point
waikiki %>%
  ggplot(aes(x = date, y = tmax)) +
  geom_point(color = "blue")
```

## Multi-panel plots

```{r}
ggp_scatter =
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_density = 
  weather_df %>%
  ggplot(aes(x = tmin)) +
  geom_density()

ggp_box = 
  weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()

ggp_scatter
ggp_density
ggp_box

# shows the two plots side by side 
ggp_scatter + (ggp_density / ggp_box)
```

## data manipulation

```{r}
weather_df %>%
  mutate(
    name = factor(name),
    name = fct_relevel(name, "Waikiki_HA", "CentralPark_NY")
  ) %>%
  ggplot(aes(x = name, y = tmax, color = name)) +
  geom_boxplot()
```

reorder instead of relevel
reoder can put things in order according to a different variable, so in this example instead of by location, it will be according to tmax (Waterhole has the lowest tmax so it continues from there)

```{r}
weather_df %>%
  mutate(
    name = factor(name),
    name = fct_reorder(name, tmax)
  ) %>%
  ggplot(aes(x = name, y = tmax, color = name)) +
  geom_boxplot()
```

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point()
```

## restructure the plot

```{r}
weather_df %>%
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperature"
    ) %>%
  ggplot(aes(x = temperature, fill = observation)) +
  geom_density(alpha = .5) +
  facet_grid(~name) +
  theme(legend.position = "bottom")
```

litters and pups

```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  mutate(wt_gain = gd18_weight - gd0_weight,
         day_of_tx = as.numeric(day_of_tx))

fas_data = left_join(pup_data, litter_data, by = "litter_number") 
```

```{r}
fas_data %>%
  ggplot(aes(x = dose, y = pd_ears)) +
  geom_violin() +
  facet_grid(day_of_tx ~.)
```

This is more of a data tidying problem and not a ggplot problem

```{r}
fas_data %>% 
  select(sex, dose, day_of_tx, pd_ears:pd_walk) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day") %>% 
  drop_na() %>% 
  mutate(outcome = forcats::fct_reorder(outcome, day_of_tx, median)) %>% 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() + 
  facet_grid(day_of_tx ~ outcome)

# facet on both day of treatment and outcome of interest 
```


